---
title: "SURV 617 Final Project Code"
output: html_document
date: "2024-12-09"
authors: "Molly Fisch-Friedman, Sean Johnson, and Brittany Owen"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3)
```

#1. PREP
##a. Packages
```{r }
library(dplyr)
library(readxl)
library(tidyverse)
library(gtsummary)
library(ggplot2)
library(lme4)
library(nlme)
library(lmerTest)
library(knitr)
library(kableExtra)
library(janitor)
library(gridExtra)
library(geepack)
```

##b. Custom Function
```{r }
lrt.pvalue <- function(fit1, fit2) {
  lrtstat <-
    (-2 * as.numeric(summary(fit2)$logLik)) - (-2 * as.numeric(summary(fit1)$logLik))
  pvalue = 0.5 * (1 - pchisq(lrtstat, 2)) + 0.5 * (1 - pchisq(lrtstat, 1))
  return(pvalue)
}
```

##c. Variables
Vars of interest

PARENT Qs:
PIMPREL - 118) P79. TEXT: Thanks, Now I have some questions about religion. How important is your religious faith in providing guidance in your own day-to-day living?: [READ LIST] (PIMPREL)

YOUTH Qs:
FAITH1 - 324) P57. TEXT: Next, I have some questions about religion. There may be some that don't directly apply to you, but just answer them as well as you can. How important or unimportant is religious faith in shaping [INSERT A-B]?: [READ LIST] How you live your daily life. (FAITH1)
ATTREG (already included) - 330) Y60. Do you attend religious services more than once or twice a year, NOT counting weddings, baptisms, and funerals [INSERT 'or religious services during school' IF teen attends a private religious school]?[DOES NOT INCLUDE RELIGIOUS YOUTH GROUP MEETINGS] (ATTREG)
PRAYALON - 670) Y230. How often, if ever, do you pray by yourself alone?[READ LIST] (PRAYALON)
READBIBL - 671) Y231. How often, if ever, do you read from [SCRIPTURES C] to yourself alone? [READ LIST] (READBIBL)

PINCOME (RECODED INCOME) - 725) Household income merged into a single scale. (PINCOME)
TEENRACE (RECODED TEEN RACE) 726) Merged and collapsed teen race. (TEENRACE)

##d. Load Data
```{r, eval=FALSE }
wave1 <- read_xlsx("~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/National Study of Youth and Religion, Wave 1 (2003).XLSX")
wave1$year <- 0

wave2 <- read_xlsx("~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/National Study of Youth and Religion, Wave 2 (2005).XLSX")
wave2$year <- 2

wave3 <- read_xlsx("~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/National Study of Youth and Religion, Wave 3 (2007-2008).XLSX")
wave3$year <- 4

wave3$NA_total <- rowSums(is.na(wave3))
tabyl(wave3$NA_total) #838 missing 465 or 570 values. 

wave4 <- read_xlsx("~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/National Study of Youth and Religion Wave 4 (2013).XLSX")
wave4$year <- 10

#names <- names(wave4)
#view(names)
```

##e. Reduce vars in files across years
```{r, eval=FALSE }
wave1_r <- wave1 %>% dplyr::select(c("IDS", "year", "RWEIGHT1", "RWEIGHT2", "NWEIGHT2", "INWAVE2", "PLIVE", "FAITH1", "PIMPREL", "PSCHREL", "PATTEND", "PDADEDUC", "PMOMEDUC", "PINCOME","EATADULT", "ETHRACE", "ATTREG", "ATTEND1", "DOUBTS1", "WISHATT1", "EDGOAL1", "EDGOAL3", "GRADES", "MEANLESS", "ABSTAIN1", "DIVCEOK", "CRRACE", "POLACT", "AFTRLIFE", "VIEWREL", "SEXEVER", "BODY", "SAD", "ACCEPTED", "GUILT1", "GUILT2", "GUILT3", "PRAYALON", "READBIBL", "GRACE", "CHWELCOM", "TEENRACE", "TEENSEX", "AGECATS", "PRACEOLD", "PWKMOMFT", "PWKMOMPT", "PWKDADFT", "PWKDADPT", "STOPATT1", "STOPATT2", "TEENNUM", "WISHATT", "XCHLDNUM", "CENREG", "CENDIV"))
wave1_r <- wave1_r %>% rename("GENDER"="TEENSEX")

#names(wave2)
wave2_r <- wave2 %>% dplyr::select(c("IDS", "year", "FAITH1", "EATADULT", "ATTREG", "ATTEND1", "DOUBTS1", "WISHATT1", "EDGOAL1", "EDGOAL3", "GRADES", "MEANLESS",   "ABSTAIN1", "AFTRLIFE", "VIEWREL", "SEXEVER", "BODY", "SAD", "ACCEPTED", "GUILT1", "PRAYALON", "READBIBL", "GRACE", "CHWELCOM", "AGECATS",                                      ))

#names(wave3)
wave3_r <- wave3 %>% dplyr::select(c("IDS", "year", "FAITH1", "ATTREG", "ATTEND1", "DOUBTS1", "MEANLESS", "ABSTAIN1", "CRRACE", "AFTRLIFE", "VIEWREL", "SEXEVER", "BODY", "SAD", "ACCEPTED", "PRAYALON", "READBIBL", "GRACE", "AGECATS", "CENDIV"))

#names(wave4)
wave4_r <- wave4 %>% dplyr::select(c("IDS", "year", "FAITH1_W4", "ATTREG_W4", "ATTEND1_W4", "DOUBTS1_W4", "MEANLESS_W4", "ABSTAIN1_W4", "VIEWREL_W4", "BODY_W4", "PRAYALON_W4", "GRACE_W4", "AGECATS_W4", "CENREG_W4", "CENDIV_W4"
  ))

wave4_r <- wave4_r %>% rename("FAITH1"="FAITH1_W4",
                              "ATTREG"="ATTREG_W4",
                              "ATTEND1"="ATTEND1_W4", 
                              "DOUBTS1"="DOUBTS1_W4", 
                              "MEANLESS"="MEANLESS_W4",
                              "ABSTAIN1"="ABSTAIN1_W4",
                              "VIEWREL"="VIEWREL_W4",
                              "BODY"="BODY_W4",
                              "PRAYALON"="PRAYALON_W4",
                              "GRACE"="GRACE_W4",
                              "AGECATS"="AGECATS_W4",
                              "CENREG"="CENREG_W4",
                              "CENDIV"="CENDIV_W4"
                              )

#DOESN'T WORK YET
#waves <- rbind(wave1_r, wave2_r, wave3_r, wave4_r)
```

##f. Impute Baseline Vars
```{r, eval=FALSE }
#BASELINE TEEN RACE
wave1$TEENRACE_bl <- wave1$TEENRACE
wave2$TEENRACE_bl <- wave1$TEENRACE[match(wave2$IDS, wave1$IDS)]
wave3$TEENRACE_bl <- wave1$TEENRACE[match(wave3$IDS, wave1$IDS)]
wave4$TEENRACE_bl <- wave1$TEENRACE[match(wave4$IDS, wave1$IDS)]

#BASELINE TEEN SEX
wave1$TEENSEX_bl <- wave1$TEENSEX
wave2$TEENSEX_bl <- wave1$TEENSEX[match(wave2$IDS, wave1$IDS)]
wave3$TEENSEX_bl <- wave1$TEENSEX[match(wave3$IDS, wave1$IDS)]
wave4$TEENSEX_bl <- wave1$TEENSEX[match(wave4$IDS, wave1$IDS)]

#AGE IS IN ALL 4 - but wave3 is missing a lot
wave1$AGECATS_bl <- wave1$AGECATS
wave2$AGECATS_bl <- wave1$AGECATS[match(wave2$IDS, wave1$IDS)]
wave3$AGECATS_bl <- wave1$AGECATS[match(wave3$IDS, wave1$IDS)]
wave4$AGECATS_bl <- wave1$AGECATS[match(wave4$IDS, wave1$IDS)]

#PIMPREL - 118) P79. TEXT: Thanks, Now I have some questions about religion. How important is your religious faith in providing guidance in your own day-to-day living?: [READ LIST] (PIMPREL)
wave1$PIMPREL_bl <- wave1$PIMPREL
wave2$PIMPREL_bl <- wave1$PIMPREL[match(wave2$IDS, wave1$IDS)]
wave3$PIMPREL_bl <- wave1$PIMPREL[match(wave3$IDS, wave1$IDS)]
wave4$PIMPREL_bl <- wave1$PIMPREL[match(wave4$IDS, wave1$IDS)]

#Baseline religion
tabyl(wave1$RELIG62)
tabyl(wave2$RELIG62)
#tabyl(wave3$RELIG62) #Not in wave
#tabyl(wave4$RELIG62) #Not in wave

wave1$RELIG62_bl <- wave1$RELIG62
wave2$RELIG62_bl <- wave1$RELIG62[match(wave2$IDS, wave1$IDS)]
wave3$RELIG62_bl <- wave1$RELIG62[match(wave3$IDS, wave1$IDS)]
wave4$RELIG62_bl <- wave1$RELIG62[match(wave4$IDS, wave1$IDS)]
```

###f. Continued
```{r, eval=FALSE }
#Baseline pray alone
wave1$PRAYALON_bl <- wave1$PRAYALON
wave2$PRAYALON_bl <- wave1$PRAYALON[match(wave2$IDS, wave1$IDS)]
wave3$PRAYALON_bl <- wave1$PRAYALON[match(wave3$IDS, wave1$IDS)]
wave4$PRAYALON_bl <- wave1$PRAYALON[match(wave4$IDS, wave1$IDS)]

wave2 %>% gtsummary::select(PRAYALON, PRAYALON_bl) %>%
  tbl_summary(by = PRAYALON,
              missing = "always")
#1 Never
#2 Less than once a month
#3 One to twice times a month
#4 Less than once a month
#5 About once a week
#6 About once a day
#7 Many times a day
#777 DK
#888 Refused

#Baseline family says grace at dinner
wave1$GRACE_bl <- wave1$GRACE
wave2$GRACE_bl <- wave1$GRACE[match(wave2$IDS, wave1$IDS)]
wave3$GRACE_bl <- wave1$GRACE[match(wave3$IDS, wave1$IDS)]
wave4$GRACE_bl <- wave1$GRACE[match(wave4$IDS, wave1$IDS)]

#Read Bible - wave 4 has different vars, not sure why?
tabyl(wave4$READBIBL1_W4)
tabyl(wave4$READBIBL2_W4)
tabyl(wave4$READBIBL3_W4)
tabyl(wave4$READBIBL2_W4)

#Read Bible - baseline
wave1$READBIBL_bl <- wave1$READBIBL
wave2$READBIBL_bl <- wave1$READBIBL[match(wave2$IDS, wave1$IDS)]
wave3$READBIBL_bl <- wave1$READBIBL[match(wave3$IDS, wave1$IDS)]
wave4$READBIBL_bl <- wave1$READBIBL[match(wave4$IDS, wave1$IDS)]

#ATTREG baseline
wave1$ATTREG_bl <- wave1$ATTREG
wave2$ATTREG_bl <- wave1$ATTREG[match(wave2$IDS, wave1$IDS)]
wave3$ATTREG_bl <- wave1$ATTREG[match(wave3$IDS, wave1$IDS)]
wave4$ATTREG_bl <- wave1$ATTREG[match(wave4$IDS, wave1$IDS)]

#RELTARD - not in all years - set baseline
wave1$RELTRAD_bl <- wave1$RELTRAD
wave2$RELTRAD_bl <- wave1$RELTRAD[match(wave2$IDS, wave1$IDS)]
wave3$RELTRAD_bl <- wave1$RELTRAD[match(wave3$IDS, wave1$IDS)]
wave4$RELTRAD_bl <- wave1$RELTRAD[match(wave4$IDS, wave1$IDS)]

tabyl(wave1$RELTRAD_bl)
tabyl(wave2$RELTRAD_bl)
tabyl(wave3$RELTRAD_bl)
tabyl(wave4$RELTRAD_bl)
```

##g. Combine Waves
```{r, eval=FALSE }
wave1_r <- wave1 %>% dplyr::select(c("IDS", "year", "ATTREG", "ATTREG_bl", "PIMPREL_bl", "RELIG62_bl", "RELTRAD_bl", "FAITH1", "PRAYALON_bl", "GRACE_bl", "READBIBL_bl", "TEENRACE_bl", "TEENSEX_bl", "AGECATS_bl", "AGECATS"))

wave2_r <- wave2 %>% dplyr::select(c("IDS", "year", "ATTREG", "ATTREG_bl", "PIMPREL_bl", "RELIG62_bl", "RELTRAD_bl", "FAITH1", "PRAYALON_bl", "GRACE_bl", "READBIBL_bl", "TEENRACE_bl", "TEENSEX_bl", "AGECATS_bl", "AGECATS"))

wave3_r <- wave3 %>% dplyr::select(c("IDS", "year","ATTREG", "ATTREG_bl", "PIMPREL_bl", "RELIG62_bl", "RELTRAD_bl", "FAITH1", "PRAYALON_bl", "GRACE_bl", "READBIBL_bl", "TEENRACE_bl", "TEENSEX_bl", "AGECATS_bl", "AGECATS"))

wave4_r <- wave4 %>% dplyr::select(c("IDS", "year","ATTREG_W4", "ATTREG_bl", "PIMPREL_bl", "RELIG62_bl", "RELTRAD_bl", "FAITH1_W4", "PRAYALON_bl", "GRACE_bl", "READBIBL_bl", "TEENRACE_bl", "TEENSEX_bl", "AGECATS_bl", "AGECATS_W4"))

wave4_r <- wave4_r %>% rename("ATTREG" = "ATTREG_W4",
                              "FAITH1" = "FAITH1_W4",
                              "AGECATS" = "AGECATS_W4")

#Combine Waves into single DF
waves <- rbind(wave1_r, wave2_r, wave3_r, wave4_r)

save(waves, file = "~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/Combined Data.RData")
```

#2. LOAD AND REVIEW PARTIALLY CLEANED DATA
```{r, eval=FALSE }
load("~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/Combined Data.RData")

waves$NA_total <- rowSums(is.na(waves))
tabyl(waves, NA_total, year) # missing 3

#11,488 total cases - 838 missing 4 in year 3 = 10,650
waves <- subset(waves, year != 3 & NA_total != 3)
tabyl(waves, NA_total, year)
```

```{r, eval=FALSE }
waves %>% gtsummary::select(year, RELIG62_bl, RELTRAD_bl, ATTREG, ATTREG_bl, PIMPREL_bl, RELIG62_bl, RELTRAD_bl, FAITH1, PRAYALON_bl, GRACE_bl, READBIBL_bl, TEENRACE_bl, TEENSEX_bl, AGECATS_bl, AGECATS) %>%
  tbl_summary(by = year,
              type = list(TEENRACE_bl ~ "categorical",
                          TEENSEX_bl ~ "categorical", 
                          AGECATS ~ "categorical",
                          RELIG62_bl ~ "categorical"),
              missing = "always")
```

##a. Cleaning
```{r, eval=FALSE }
#Set Refuse/DK/etc to NA
tabyl(waves$RELIG62_bl)
tabyl(waves$RELTRAD_bl)

tabyl(waves$ATTREG)
waves$ATTREG[waves$ATTREG == 666] <- NA
waves$ATTREG[waves$ATTREG == 777] <- NA
waves$ATTREG[waves$ATTREG == 888] <- NA
tabyl(waves$ATTREG)

tabyl(waves$ATTREG_bl)
waves$ATTREG_bl[waves$ATTREG_bl == 777] <- NA
waves$ATTREG_bl[waves$ATTREG_bl == 888] <- NA
tabyl(waves$ATTREG_bl)

tabyl(waves$PIMPREL_bl)
waves$PIMPREL_bl[waves$PIMPREL_bl == 777] <- NA
waves$PIMPREL_bl[waves$PIMPREL_bl == 888] <- NA
tabyl(waves$PIMPREL_bl)

tabyl(waves$RELIG62_bl)
waves$RELIG62_bl[waves$RELIG62_bl == 777] <- NA
waves$RELIG62_bl[waves$RELIG62_bl == 888] <- NA
tabyl(waves$RELIG62_bl)

tabyl(waves$FAITH1)
waves$FAITH1[waves$FAITH1 == 666] <- NA
waves$FAITH1[waves$FAITH1 == 777] <- NA
waves$FAITH1[waves$FAITH1 == 888] <- NA
tabyl(waves$FAITH1)

tabyl(waves$PRAYALON_bl)
waves$PRAYALON_bl[waves$PRAYALON_bl == 777] <- NA
waves$PRAYALON_bl[waves$PRAYALON_bl == 888] <- NA
tabyl(waves$PRAYALON_bl)

tabyl(waves$GRACE_bl)
waves$GRACE_bl[waves$GRACE_bl == 777] <- NA
waves$GRACE_bl[waves$GRACE_bl == 888] <- NA
tabyl(waves$GRACE_bl)

tabyl(waves$READBIBL_bl)
waves$READBIBL_bl[waves$READBIBL_bl == 777] <- NA
waves$READBIBL_bl[waves$READBIBL_bl == 888] <- NA
tabyl(waves$READBIBL_bl)

tabyl(waves$TEENRACE_bl)
waves$TEENRACE_bl[waves$TEENRACE_bl == 777] <- NA
waves$TEENRACE_bl[waves$TEENRACE_bl == 888] <- NA
tabyl(waves$TEENRACE_bl)

tabyl(waves$AGECATS_bl)
waves$AGECATS_bl[waves$AGECATS_bl == 888] <- NA
tabyl(waves$AGECATS_bl)

tabyl(waves$AGECATS)
waves$AGECATS[waves$AGECATS == 888] <- NA
tabyl(waves$AGECATS)
```

##b. Review Vars (again)
```{r, eval=FALSE }
waves %>% gtsummary::select(year, RELIG62_bl, RELTRAD_bl, ATTREG, ATTREG_bl, PIMPREL_bl, RELIG62_bl, RELTRAD_bl, FAITH1, PRAYALON_bl, GRACE_bl, READBIBL_bl, TEENRACE_bl, TEENSEX_bl, AGECATS_bl, AGECATS) %>%
  tbl_summary(by = year,
              type = list(ATTREG ~ "categorical",
                          ATTREG_bl ~ "categorical",
                          GRACE_bl ~ "categorical",
                          TEENRACE_bl ~ "categorical",
                          TEENSEX_bl ~ "categorical", 
                          AGECATS ~ "categorical",
                          RELIG62_bl ~ "categorical"),
              missing = "always")
```

##c. Review Cases with missing data
```{r, eval=FALSE }
waves$NA_total <- rowSums(is.na(waves))

tabyl(waves$NA_total) #97.3% have no missing data
tabyl(waves, NA_total, year) 
```

##d. Recodes
```{r, eval=FALSE }
tabyl(waves$RELIG62_bl)
waves$RELIG62_bl_r[waves$RELIG62_bl == 1 ] <- "1 Catholic"
waves$RELIG62_bl_r[waves$RELIG62_bl == 2 ] <- "2 Protestant"
waves$RELIG62_bl_r[waves$RELIG62_bl == 3 ] <- "2 Protestant"
waves$RELIG62_bl_r[waves$RELIG62_bl == 4 ] <- "3 Jewish"
waves$RELIG62_bl_r[waves$RELIG62_bl == 5 ] <- "4 Other"
waves$RELIG62_bl_r[waves$RELIG62_bl == 6 ] <- "4 Other"
waves$RELIG62_bl_r[waves$RELIG62_bl == 7 ] <- "4 Other"
waves$RELIG62_bl_r[waves$RELIG62_bl == 8 ] <- "5 Not religious"
tabyl(waves$RELIG62_bl_r)

waves$RELIG62_bl[waves$RELIG62_bl == 1 ] <- "1 Catholic"
waves$RELIG62_bl[waves$RELIG62_bl == 2 ] <- "2 Protestant"
waves$RELIG62_bl[waves$RELIG62_bl == 3 ] <- "3 Christian"
waves$RELIG62_bl[waves$RELIG62_bl == 4 ] <- "4 Jewish"
waves$RELIG62_bl[waves$RELIG62_bl == 5 ] <- "5 Muslim/Islamic"
waves$RELIG62_bl[waves$RELIG62_bl == 6 ] <- "6 Another religion"
waves$RELIG62_bl[waves$RELIG62_bl == 7 ] <- "7 Half one religiion and hafl another"
waves$RELIG62_bl[waves$RELIG62_bl == 8 ] <- "8 Not religious"
tabyl(waves$RELIG62_bl)

tabyl(waves$RELTRAD_bl)
waves$RELTRAD_bl[waves$RELTRAD_bl == 1 ] <- "1 Conservative Protestant"
waves$RELTRAD_bl[waves$RELTRAD_bl == 2 ] <- "2 Mainline Protestant"
waves$RELTRAD_bl[waves$RELTRAD_bl == 3 ] <- "3 Black Protestant"
waves$RELTRAD_bl[waves$RELTRAD_bl == 4 ] <- "4 Catholic"
waves$RELTRAD_bl[waves$RELTRAD_bl == 5 ] <- "5 Jewish"
waves$RELTRAD_bl[waves$RELTRAD_bl == 6 ] <- "6 Mormon/LDS"
waves$RELTRAD_bl[waves$RELTRAD_bl == 7 ] <- "7 Unaffiliated "
waves$RELTRAD_bl[waves$RELTRAD_bl == 8 ] <- "8 Other religion"
waves$RELTRAD_bl[waves$RELTRAD_bl == 9 ] <- "9 Indeterminate"
tabyl(waves$RELTRAD_bl)

tabyl(waves$ATTREG) #1 = yes
tabyl(waves$ATTREG_bl) #1 = yes 

tabyl(waves$PIMPREL_bl)
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 1] <- "1 Extremely or very important"
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 2] <- "1 Extremely or very important"
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 3] <- "2 Fairly or somewhat important"
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 4] <- "2 Fairly or somewhat important"
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 5] <- "3 Not very or not at all important"
waves$PIMPREL_bl_r[waves$PIMPREL_bl == 6] <- "3 Not very or not at all important"
tabyl(waves$PIMPREL_bl_r)

tabyl(waves$PIMPREL_bl)
waves$PIMPREL_bl[waves$PIMPREL_bl == 1 ] <- "1 Extremely important"
waves$PIMPREL_bl[waves$PIMPREL_bl == 2 ] <- "2 Very important"
waves$PIMPREL_bl[waves$PIMPREL_bl == 3 ] <- "3 Fairly important"
waves$PIMPREL_bl[waves$PIMPREL_bl == 4 ] <- "4 Somewhat important"
waves$PIMPREL_bl[waves$PIMPREL_bl == 5 ] <- "5 Not very important"
waves$PIMPREL_bl[waves$PIMPREL_bl == 6 ] <- "6 Not important at all"
tabyl(waves$PIMPREL_bl)

tabyl(waves$GRACE_bl)#1 = yes
waves$GRACE_bl[waves$GRACE_bl == 1 ] <- "1 Yes, say grace"
waves$GRACE_bl[waves$GRACE_bl == 0 ] <- "2 No, don't say grace"
tabyl(waves$GRACE_bl)

tabyl(waves$READBIBL_bl)
#Set binary for models
waves$READBIBL_bl_binary[waves$READBIBL_bl == 1] <- "2 Never read text" #never
waves$READBIBL_bl_binary[waves$READBIBL_bl != 1] <- "1 Read text"
tabyl(waves$READBIBL_bl_binary)

waves$READBIBL_bl[waves$READBIBL_bl == 1 ] <- "7 Never"
waves$READBIBL_bl[waves$READBIBL_bl == 2 ] <- "6 Less than once a month"
waves$READBIBL_bl[waves$READBIBL_bl == 3 ] <- "5 One to two times a month"
waves$READBIBL_bl[waves$READBIBL_bl == 4 ] <- "4 About once a week"
waves$READBIBL_bl[waves$READBIBL_bl == 5 ] <- "3 A few times a week"
waves$READBIBL_bl[waves$READBIBL_bl == 6 ] <- "2 About once a day"
waves$READBIBL_bl[waves$READBIBL_bl == 7 ] <- "1 Many times a day"
tabyl(waves$READBIBL_bl)

tabyl(waves$FAITH1)
waves$FAITH1_r[waves$FAITH1 == 1] <- "1 Extremely or very important"
waves$FAITH1_r[waves$FAITH1 == 2] <- "1 Extremely or very important"
waves$FAITH1_r[waves$FAITH1 == 3] <- "2 Somewhat important"
waves$FAITH1_r[waves$FAITH1 == 4] <- "3 Not very or not at all important"
waves$FAITH1_r[waves$FAITH1 == 5] <- "3 Not very or not at all important"
tabyl(waves$FAITH1_r)

tabyl(waves$FAITH1)
waves$FAITH1[waves$FAITH1 == 1] <- "1 Extremely important"
waves$FAITH1[waves$FAITH1 == 2] <- "2 Very important"
waves$FAITH1[waves$FAITH1 == 3] <- "3 Somewhat important"
waves$FAITH1[waves$FAITH1 == 4] <- "4 Not very important"
waves$FAITH1[waves$FAITH1 == 5] <- "5 Not important at all"
tabyl(waves$FAITH1)

tabyl(waves$PRAYALON_bl)
waves$PRAYALON_bl_binary[waves$PRAYALON_bl == 1] <- "2 Never pray"
waves$PRAYALON_bl_binary[waves$PRAYALON_bl != 1] <- "1 Prays"
tabyl(waves$PRAYALON_bl_binary)

waves$PRAYALON_bl[waves$PRAYALON_bl == 1] <- "7 Never"
waves$PRAYALON_bl[waves$PRAYALON_bl == 2] <- "6 Less than once a month"
waves$PRAYALON_bl[waves$PRAYALON_bl == 3] <- "5 One to two times a month"
waves$PRAYALON_bl[waves$PRAYALON_bl == 4] <- "4 About once a week"
waves$PRAYALON_bl[waves$PRAYALON_bl == 5] <- "3 A few times a week"
waves$PRAYALON_bl[waves$PRAYALON_bl == 6] <- "2 About once a day"
waves$PRAYALON_bl[waves$PRAYALON_bl == 7] <- "1 Many times a day"
tabyl(waves$PRAYALON_bl)

tabyl(waves$TEENRACE_bl)
#White binary
waves$TEENRACE_bl_white[waves$TEENRACE_bl == 1]  <- "1 White" #Yes, white
waves$TEENRACE_bl_white[waves$TEENRACE_bl != 1]  <- "2 Not white"
tabyl(waves$TEENRACE_bl_white)

tabyl(waves$TEENRACE_bl)
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 1] <- "1 White"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 2] <- "2 Black"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 3] <- "3 Hispanic"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 4] <- "4 Mixed or other"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 5] <- "4 Mixed or other"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 6] <- "4 Mixed or other"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 7] <- "4 Mixed or other"
waves$TEENRACE_bl_r[waves$TEENRACE_bl == 8] <- "4 Mixed or other"
tabyl(waves$TEENRACE_bl_r)

waves$TEENRACE_bl[waves$TEENRACE_bl == 1]  <- "1 White"
waves$TEENRACE_bl[waves$TEENRACE_bl == 2]  <- "2 Black"
waves$TEENRACE_bl[waves$TEENRACE_bl == 3]  <- "3 Hispanic"
waves$TEENRACE_bl[waves$TEENRACE_bl == 4]  <- "4 Asian"
waves$TEENRACE_bl[waves$TEENRACE_bl == 5]  <- "5 Islander"
waves$TEENRACE_bl[waves$TEENRACE_bl == 6]  <- "6 Native American"
waves$TEENRACE_bl[waves$TEENRACE_bl == 7]  <- "7 Mixed"
waves$TEENRACE_bl[waves$TEENRACE_bl == 8]  <- "8 Other"
tabyl(waves$TEENRACE_bl)

tabyl(waves$TEENSEX_bl)
waves$TEENSEX_bl[waves$TEENSEX_bl == 0]  <- "1 Male"
waves$TEENSEX_bl[waves$TEENSEX_bl == 1]  <- "2 Female"
tabyl(waves$TEENSEX_bl)

tabyl(waves$AGECATS)

#20+ binary
waves$AGECATS_20plus[waves$AGECATS >= 20]  <- "1 20+"
waves$AGECATS_20plus[waves$AGECATS < 20]  <- "2 Under 20"
tabyl(waves$AGECATS_20plus) #37.5% above 20
```

```{r, eval=FALSE }
waves %>% gtsummary::select(year, RELIG62_bl, RELTRAD_bl, ATTREG, ATTREG_bl, PIMPREL_bl, PIMPREL_bl_r, RELIG62_bl, RELIG62_bl_r, RELTRAD_bl, FAITH1, FAITH1_r, PRAYALON_bl, PRAYALON_bl_binary, GRACE_bl, READBIBL_bl, READBIBL_bl_binary, TEENRACE_bl, TEENRACE_bl_r, TEENRACE_bl_white, TEENSEX_bl, AGECATS_bl, AGECATS, AGECATS_20plus) %>%
  tbl_summary(by = year,
              type = list(ATTREG ~ "categorical",
                          ATTREG_bl ~ "categorical",
                          GRACE_bl ~ "categorical",
                          TEENRACE_bl ~ "categorical",
                          TEENSEX_bl ~ "categorical", 
                          AGECATS ~ "categorical",
                          RELIG62_bl ~ "categorical"),
              missing = "always")
```

##e. Save fully cleaned data
```{r, eval=FALSE }
save(waves, file = "~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/Combined Data_clean.RData")

tabyl(waves$NA_total) #97.34 not missing any data
tabyl(waves, NA_total, year) 

tabyl(waves$ATTREG_bl)
waves_attreg_all <- subset(waves, ATTREG_bl == 1)
tabyl(waves_attreg_all$NA_total) #97.67 not missing any data
tabyl(waves_attreg_all, NA_total, year) 

waves_attreg_nomissing_all <- subset(waves_attreg_all, NA_total == 0)
tabyl(waves_attreg_nomissing_all$NA_total) #97.67 not missing any data
tabyl(waves_attreg_nomissing_all, NA_total, year) 
save(waves_attreg_nomissing_all, file = "~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/Combined Data_clean_attreg1_all.RData")

tabyl(waves_attreg_nomissing_all$year)
waves_attreg_nomissing_234 <- subset(waves_attreg_nomissing_all, year != 0)
tabyl(waves_attreg_nomissing_234$year) #97.67 not missing any data
save(waves_attreg_nomissing_234, file = "~/Desktop/JPSM - Masters Program/JPSM Classes/03_Fall 2024/SURV 617 - Statistical Applications/Group Project/NSYR data/Combined Data_clean_attreg1_234.RData")
```

##f. Review vars, attreg=1, no missing data
```{r, eval=FALSE }
waves_attreg_nomissing_all %>% gtsummary::select(year, RELIG62_bl, RELTRAD_bl, ATTREG, ATTREG_bl, PIMPREL_bl, PIMPREL_bl_r, RELIG62_bl, RELIG62_bl_r, RELTRAD_bl, FAITH1, FAITH1_r, PRAYALON_bl, PRAYALON_bl_binary, GRACE_bl, READBIBL_bl, READBIBL_bl_binary, TEENRACE_bl, TEENRACE_bl_r, TEENRACE_bl_white, TEENSEX_bl, AGECATS_bl, AGECATS, AGECATS_20plus) %>%
  tbl_summary(by = year,
              type = list(ATTREG ~ "categorical",
                          ATTREG_bl ~ "categorical",
                          GRACE_bl ~ "categorical",
                          TEENRACE_bl ~ "categorical",
                          TEENSEX_bl ~ "categorical", 
                          AGECATS ~ "categorical",
                          RELIG62_bl ~ "categorical"),
              missing = "always")
```

#3. FULLY CLEANED DATA

##a. Load Fully Cleaned Data
```{r }
load("~/Downloads/Combined Data_clean.RData")

load("~/Downloads/Combined Data_clean_attreg1_all.RData")

load("~/Downloads/Combined Data_clean_attreg1_234.RData")
```

##b. Review data
```{r }
waves %>% gtsummary::select(year, RELIG62_bl, RELIG62_bl_r, RELTRAD_bl, ATTREG, ATTREG_bl, PIMPREL_bl, PIMPREL_bl_r, RELIG62_bl, RELTRAD_bl, FAITH1, FAITH1_r, PRAYALON_bl, PRAYALON_bl_binary, GRACE_bl, READBIBL_bl, READBIBL_bl_binary, TEENRACE_bl, TEENRACE_bl_r, TEENRACE_bl_white, TEENSEX_bl, AGECATS_bl, AGECATS, AGECATS_20plus) %>%
  tbl_summary(by = year,
              type = list(ATTREG ~ "categorical",
                          ATTREG_bl ~ "categorical",
                          GRACE_bl ~ "categorical",
                          PRAYALON_bl_binary ~ "categorical",
                          TEENRACE_bl_white ~ "categorical",
                          READBIBL_bl_binary ~ "categorical",
                          AGECATS ~ "categorical",
                          AGECATS_20plus ~ "categorical"),
              missing = "always")

#Doubts1 has lots of missing due to display condition
```


#5. MODELS only when ATTREG_bl=1 and 

```{r }
data <- waves_attreg_nomissing_234
table(data$year, data$ATTREG)

marginal_model <- geeglm(ATTREG ~ year,
                   data=data,
                    id=IDS,
                 family=binomial("logit"),
                 corstr = "exchangeable",
                 na.action=na.omit)

summary(marginal_model)
marginal_model %>% tbl_regression()
marginal_model %>% tbl_regression(exponentiate = T)

marginal_all <- geeglm(ATTREG ~ year + RELIG62_bl_r + year*RELIG62_bl_r +
                           TEENRACE_bl_white + year*TEENRACE_bl_white + 
                 TEENSEX_bl + year*TEENSEX_bl + factor(AGECATS_bl) + year*factor(AGECATS_bl) +
                   PIMPREL_bl_r + 
                   FAITH1_r + PRAYALON_bl_binary + GRACE_bl +
                   READBIBL_bl_binary + 
                    PIMPREL_bl_r*year + 
                   FAITH1_r*year + PRAYALON_bl_binary*year +
                     GRACE_bl*year +
                   READBIBL_bl_binary*year,
                 data=data,
                    id=IDS,
                 family=binomial("logit"),
                 corstr = "exchangeable",
                 na.action=na.omit)
summary(marginal_all)
marginal_all %>% tbl_regression()
marginal_all %>% tbl_regression(exponentiate = T)
QIC(marginal_all)

marginal_sig <- geeglm(ATTREG ~ year + RELIG62_bl_r + 
                           TEENRACE_bl_white +
                         TEENSEX_bl + factor(AGECATS_bl) +
                   PIMPREL_bl_r + 
                   FAITH1_r + GRACE_bl +
                   READBIBL_bl_binary + 
                   FAITH1_r*year,
                 data=data,
                    id=IDS,
                 family=binomial("logit"),
                 corstr = "exchangeable",
                 na.action=na.omit)
summary(marginal_sig)
marginal_sig %>% tbl_regression()
marginal_sig %>% tbl_regression(exponentiate = T)
QIC(marginal_sig)

marginal_sig_wo_grace_read <- geeglm(ATTREG ~ year + RELIG62_bl_r + 
                           TEENRACE_bl_white +
                         TEENSEX_bl + factor(AGECATS_bl) +
                   PIMPREL_bl_r + 
                   FAITH1_r +
                   FAITH1_r*year,
                 data=data,
                    id=IDS,
                 family=binomial("logit"),
                 corstr = "exchangeable",
                 na.action=na.omit)
summary(marginal_sig_wo_grace_read)
marginal_sig_wo_grace_read %>% tbl_regression()
marginal_sig_wo_grace_read %>% tbl_regression(exponentiate = T)
QIC(marginal_sig_wo_grace_read)
```


